The critical micelle concentration (CMC) of a surfactant defines the concentration above which the surfactant monomers self-assemble in solution to form micelles. It is an important property because the formation of micelles affects many interfacial phenomena \cite{rosenSurfactantsInterfacialPhenomena2012} micelles can encapsulate hydrophobic molecules to form useful complexes, but the formation of micelles also inhibits certain processes, like the aggregation of surfactants at interfaces in order to reduce surface tension.

Perhaps the most well-established predictor for CMC, $X_{cmc}$, is the Stauff-Klevens relationship, first published in \citeyear{klevensStructureAggregationDilate1953} \cite{klevensStructureAggregationDilate1953}. It formalised the observation that CMC decreases exponentially with an increase in the number of carbons in the hydrocarbon tail, $n_c$:

\begin{equation}
    \label{eq:klevens}
    \log X_{cmc} = A - Bn_c, \quad B > 0
\end{equation}

where $A$ and $B$ are empirical constants that depend on the temperature and the homologous series, i.e. the headgroup. The model is simple yet accurate, and it is easily interpretable: to reduce CMC, it is sufficient to extend the surfactant's hydrocarbon tail thus defining an easy-to-apply qualitative heuristic. Its drawback as a predictive model is its very limited applicability domain; each set of parameters is only applicable to surfactants with a specific headgroup and a linear carbon tail.
One of the goals of the quantitative structure-property relationship (QSPR) development is to produce models that are general so that we can apply them to a diverse range of compounds, design novel molecules with target properties, and
interpret the models' results to glean chemical insights.

There have been a wealth of investigations into making more general models for
CMC prediction; here, a brief review of some diverse and promising approaches
for predicting CMCs of aqueous, single-surfactant systems will be given.

\input{sections/background.tex}

Here, we build upon this previously published GNN model in two ways: we apply a
hyperparameter search algorithm to further optimise the model's architecture and
improve its accuracy, and we implement an \emph{uncertainty quantification}
technique that yields confidence intervals alongside CMC predictions. This
improved model is compared against an adaptation of the group contribution
approach that determines the entire set of groups to consider using a feature
selection routine. In addition, a separate dataset is introduced so as to
perform external validation and to probe the limits of the applicability domain
of the models, as well as to analyse the uncertainty quantification. Finally, we
discuss methods to interpret both models, and demonstrate a technique that
allows one to visualise chemical space through the `eyes' of the trained GNN. We
show that this technique can help to probe the limits of the model's
applicability domain.